---
title: "mars"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mars}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Fitting Multivariate Adapative Regression Spline (MARS) Models

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Getting started

* Install the package from GitHub. First install and load the 'devtools' package and then install 'mars' with
```{r eval = FALSE}
install_github(repo = "https://github.com/VictorWangZihe/Mars-project.git")
```

* In case that the code above is not working. The one who already have the package can try code as following.
```{r}
library(devtools)
load_all()
```

* Now load the package
```{r}
library(mars)
```


## The MARS algorithm

The MARS algorithm is a form of regression analysis introduced by Jerome H. Friedman in 1991. It is an extension of a common linear regression algorithm that takes care of nonlinearities and interactions between different variables.
Like the common linear regression, we need to specify one response variable and a set of explanatory variables to make the model.
Unlike the common linear regression, we will use the hinge function to count the nonlinear effect and hence produce a more actually statistical model.
Within the algorithm, we will use fwd_stepwise selection and bwd_stepwise selection to decide when and which variable should apply the hinge function.

## Preparing inputs

There are two required and one optional input.
1. An R formula is required to specify the response and explanatory variables.
2. A data frame containing the data to analyze is required. For this example we will use the Wage data from the ISLR package.
```{r}
library(ISLR);data(Wage)
myformula <- formula(wage ~ year + age + education)
```
3. The optional argument is a `mars.control` object. Users should use the constructor `mar.control()` to specify the three model fitting parameters.
 * The parameter `Mmax` is final number of basis function. the default value is 2.
 * The parameter `d` is the smoothing parameter the default value is 3.
 * The parameter `trace` decides whether going to trace the algorithmn. the default value is `FALSE`
```{r}
mc <- mars.control(Mmax = 6)
```
 
## Calling `mars()`
```{r}
mm <- mars(myformula,Wage,mc)
```

## Using mars methods
 * Printing a `mars` object
 * The print method for mars is very like to the common print method. It will print out the coefficient of the mars object.
 * An example output is at the end of this document.
```{r eval=FALSE}
print.mars(mm)
```
 * Get a summary
 * The summary include a set of useful information of MARS object.
 * Including coefficients, formula, variables, and Basis functions.
 * An example output is at the end of this document.
```{r eval=FALSE}
summary.mars(mm)
```
 * Get a plot
 * The plot method takes what returns by `mars()` and represents it in a visual way.
 * An example output is at the end of this document.
```{r eval=FALSE}
plot.mars(mm)
```
 * Make predictions
 * Unlike the above three methods, the `predict()` method in mars can take another variable `newdata`. It should be a set of data we want to predict.
 * If you do not have an available set of new data, the `predict()` method will predict the old data that is available. 
 * An example output is at the end of this document.
```{r eval=FALSE}
predict.mars(mm)
```
 
 
## Reference
 * Jerome H. Friedman. "Multivariate Adaptive Regression Splines". The Annals of Statistics , Mar., 1991, Vol. 19, No. 1 (Mar., 1991), pp. 1-67


## Example OutPut
Below is the example outputs for all four methods.
```{r}
print.mars(mm)
```

```{r}
summary.mars(mm)
```

```{r}
plot.mars(mm)
```

* For prediction, it should return a set of data.
* The document only show first 50 entries of that data.
```{r}
pm <- predict.mars(mm)
head(pm, n = 50L)
```




